---
title: "Codebooks"
format: gfm
---

```{r}
#| echo: false
#| eval: true
#| message: false
#| warning: false

library(haven)
library(tidyverse)
```

## Purpose

This markdown document contains codebooks for the variables constructed from raw variables in the 1930 census. The codebooks provide a description of the variables, their values, and the construction process. Should you wish to replicate the analysis or use the variables in your own analysis, the codebooks will guide you through the process.

## Variables constructed from the 1930 Swedish census

### Marital Status

The 1930 Swedish census contains a variable `civ` that captures the civil status of individuals. The variable `civ` is a string, containing values "O", "G", "E", "X", and other values. The variable `civ` is recoded into a new variable `marital_status` that captures the marital status in a more interpretable format. The recoding is done as follows:

```{stata}
#| eval: false

* Load your dataset
use "swedish_census_1930.dta", clear

* Generate the new variable marital_status
gen marital_status = ""

* Assign values to marital_status based on civ
replace marital_status = "Unmarried" if civ == "O"
replace marital_status = "Married" if civ == "G"
replace marital_status = "Widow/Widower" if civ == "E"
replace marital_status = "Divorced" if civ == "X"
replace marital_status = "Other" if !inlist(civ, "O", "G", "E", "X")

* Verify the changes
list civ marital_status
```


### Schooling

The 1930 Swedish census contains a variable `skola` that captures the level of schooling of individuals. The variable `skola` is a string, containing values from 0 to 9, as well as "X". The variable `skola` is recoded into a new variable `schooling_abbreviated` that captures the level of schooling in a more interpretable format. The recoding is done as follows:

```{stata}
#| eval: false

* Load your dataset
use "swedish_census_1930.dta", clear

* Generate the new variable schooling_abbreviated
gen schooling_abbreviated = ""

* Assign values to schooling_abbreviated based on schooling_raw
replace schooling_abbreviated = "Post-secondary and university" if inlist(schooling_raw, 8, 9)
replace schooling_abbreviated = "Post primary schooling" if inlist(schooling_raw, 4, 5, 6, 7)
replace schooling_abbreviated = "Primary school" if schooling_raw == 3
replace schooling_abbreviated = "Literate" if inlist(schooling_raw, 1, 2)
replace schooling_abbreviated = "NA" if schooling_raw == 0

* Verify the changes
list schooling_raw schooling_abbreviated
```


### HISCLASS

Occupations in 1930 have been coded into an abbreviated HISCLASS scheme in order to analyze the social class of individuals. The HISCLASS scheme is based on the HISCO classification system, which is a historical international standard classification of occupations. The HISCLASS scheme divides occupations into twelve social classes, ranging from "1. Higher managers" to "12. Unskilled farmworkers". Our abbreviated version creates seven groups from these 12 classes.


| Number | Title                                      | Number | Title                    |
|--------|--------------------------------------------|--------|--------------------------|
| 1      | Higher managers                            | 1      | Elite                    |
| 2      | Higher professionals                       |        |                          |
| 3      | Lower managers                             | 2      | White collar             |
| 4      | Lower professionals, clerical and sales personnel |        |                          |
| 5      | Lower clerical and sales personnel         |        |                          |
| 6      | Foremen                                    | 3      | Foremen                  |
| 7      | Medium-skilled workers                     | 4      | Medium-skilled workers   |
| 8      | Farmers and fishermen                      | 5      | Farmers and fishermen    |
| 9      | Low-skilled workers                        | 6      | Low-skilled workers      |
| 10     | Low-skilled farm workers                   |        |                          |
| 11     | Unskilled workers                          | 7      | Unskilled workers        |
| 12     | Unskilled farm workers                     |        |                          |


Because the 1930 census lacked hisco codes associated with the occupational titles when we got access to it, we used the hisco codes from the 1900 and 1910 censuses from IPUMS to assign hisco codes to the 1930 census. The hisco codes were then used to assign the HISCLASS groups to the 1930 census. For the common occupational titles that were included in the 1930 census but did not exist in the 1900 and 1910 censuses, we hand coded the hisco codes and assigned the HISCLASS groups. The complete list of occupational titles and their associated HISCO codes and HISCLASS groups can be found in the dataset `hisco_1930.dta`, in the codebooks folder.

The dataset `hisco_1930.dta` contains the following variables:

- `occupation_title`: The title of the occupation.
- `hisco_code`: The HISCO code associated with the occupation.
- `hisclass_group`: The HISCLASS group assigned to the occupation.

An excerpt of the dataset is shown below:

```{r}
hisco_1930 <- read_dta("hisco_1930.dta")

hisco_1930 %>% head()
```

### Electricity job (direct)

We define a new variable `electricity_job_direct` that captures whether an individual has a job directly related to electricity. The variable `electricity_job_direct` is a binary variable, taking the value 1 if the individual has a job directly related to electricity and 0 otherwise. The variable `electricity_job_direct` is constructed based on the variable `yrke`, which contains the occupation of the individual. The variable `electricity_job_direct` is constructed as follows:

```{stata}
#| eval: false

* Load your dataset
use "swedish_census_1930.dta", clear

* Generate the new variable electricity_job_direct
gen electricity_job_direct = 0

* Assign values to electricity_job_direct based on yrke
replace electricity_job_direct = 1 if regexm(yrke, "elektr", "i")
replace electricity_job_direct = 1 if yrke == "Elmontör"
replace electricity_job_direct = 1 if yrke == "Linjearbetare"
replace electricity_job_direct = 1 if yrke == "Kraftverksarbetare"

* Verify the changes
list yrke electricity_job_direct
```

### Electricity job (indirect)

Similarly, we create a variable called `electricity_job_indirect` that captures whether an individual has a job indirectly related to electricity. The variable `electricity_job_indirect` is a binary variable, taking the value 1 if the individual has a job indirectly related to electricity and 0 otherwise. The variable `electricity_job_indirect` is constructed based on the variable `yrke`, which contains the occupation of the individual. The variable `electricity_job_indirect` is hand coded from a list of common occupations in the 1930 census that are indirectly related to electricity. 

The list of indirect electricity jobs is located at `indirect_electricity_jobs_1930.dta`, in the codebooks folder. The dataset `indirect_electricity_jobs_1930.dta` contains the following variables:

- `occupation`: The occupation that is indirectly related to electricity.

An excerpt of the dataset is shown below:

```{r}
indirect_electricity_jobs_1930 <- read_dta("indirect_electricity_jobs_1930.dta")

indirect_electricity_jobs_1930 %>% head()
```

To code this up in Stata, you can use the following code:

```{stata}
#| eval: false

* Load the list of indirect electricity jobs
use "codebooks/indirect_electricity_jobs_1930.dta", clear

* Create a local macro with the list of occupations
local occupations
levelsof occupation, local(occupations)

* Load your main dataset again
use "swedish_census_1930.dta", clear

* Generate the new variable indirect_electricity_job
gen indirect_electricity_job = 0

* Assign values to indirect_electricity_job based on yrke
foreach occ in `occupations' {
    replace indirect_electricity_job = 1 if yrke == "`occ'"
}

* Verify the changes
list yrke indirect_electricity_job
```

### Employed

The variable `employed` is a binary variable, taking the value 1 if the individual is employed and 0 otherwise. The variable `employed` is constructed based on the variable `hisclass_group`, which contains the HISCLASS group of the individual. The variable `employed` is constructed as follows:

```{stata}
#| eval: false

* Load your dataset
use "swedish_census_1930.dta", clear

* Generate the new variable employed
gen employed = 0

* Assign value 1 to employed if hisclass_group is not missing
replace employed = 1 if !missing(hisclass_group)

* Verify the changes
list hisclass_group employed
```

### Union Membership

```{stata}
#| eval: false

* Load the data from the Excel file
use "union_density_1930.dta", clear

* Group by 'parish_code' and 'type_of_organization' and sum the number of members for 1900, 1910, and 1930
collapse (sum) n_members_1900 n_members_1910 n_members_1930 ///
         (first) population_1900 population_1910 population_1930, ///
         by(parish_code type_of_organization)

* Calculate union density for 1900, 1910, and 1930
gen union_density_1900 = (n_members_1900 / population_1900) * 100
gen union_density_1910 = (n_members_1910 / population_1910) * 100
gen union_density_1930 = (n_members_1930 / population_1930) * 100

* Load the original data again to get 'western_line' and 'distance_to_line'
use "union_density_1930.dta"

* Merge the grouped data with the original data
merge 1:m parish_code using "data/parishes/union_membership_and_western_line_groups_merged.xlsx"

* Drop duplicates
duplicates drop

* Group parishes into 'western_line', 'control', and 'other' based on the conditions
gen group = "other"
replace group = "western_line" if western_line == 1
replace group = "control" if western_line == 0 & distance_to_line < 300

* Cap the union densities at 100% for all years
replace union_density_1900 = min(union_density_1900, 100)
replace union_density_1910 = min(union_density_1910, 100)
replace union_density_1930 = min(union_density_1930, 100)

* Pivot the data
reshape wide union_density_1900 union_density_1910 union_density_1930, i(parish_code) j(type_of_organization) string

* Replace missing values with zero
foreach var of varlist union_density_* {
    replace `var' = 0 if missing(`var')
}

* Verify the changes
list parish_code union_density_1900 union_density_1910 union_density_1930
```

### Distance to the Western Line

```{r}
#| eval: false

library(readxl)
library(sf)
library(dplyr)

run_analysis <- function(distances, source_type, outcome_var, exclude_touching, include_area, include_population) {  
  # Reading shapefile
  parishes_path <- 'data/maps/Swedish_parishes_1926.shp'
  parishes <- st_read(parishes_path)
  # Ensure both sf objects use the same Coordinate Reference System
  parishes <- st_transform(parishes, crs = 4326)
  
  # Convert CRS to UTM zone 33N
  parishes <- st_transform(parishes, crs = 32633)
  # Calculate area in square kilometers and add as a new column
  parishes$area_sqkm <- st_area(parishes) / 10^6
  # Convert CRS back to latitude and longitude
  parishes <- st_transform(parishes, crs = 4326)
  # Create a new column called ref_code_short by removing the last 3 characters from ref_code
  population_df$ref_code_short <- substr(population_df$ref_code, 1, nchar(population_df$ref_code) - 3)
  # Drop ref_code column
  population_df <- population_df %>% select(-ref_code)
  # Do the same for parishes
  parishes$ref_code_short <- substr(parishes$ref_code, 1, nchar(parishes$ref_code) - 3)
  # Merge population_1930 to parishes
  parishes <- left_join(parishes, population_df, by = "ref_code_short")
  
  # Spatial join
  stations_with_parish <- st_join(geo_power_stations, parishes, join = st_within)
  num_power_stations <- stations_with_parish %>% group_by(ref_code) %>% summarise(num_power_stations = n())
  
  # Create points for Power Station
  station1_coords <- st_point(c(12.272571287279908, 58.27550603306433))
  station2_coords <- st_point(c(17.443262677788724, 60.5633613884066))
  
  # Create line
  line <- st_sfc(st_linestring(rbind(station1_coords, station2_coords)), crs = 4326)
  
  # Create a GeoDataFrame from the line
  line_gdf <- st_sf(geometry = line)
  
  # Create buffers
  buffer_gdfs <- lapply(distances, function(distance) {
    st_transform(line_gdf, crs = 32633) %>%
      st_buffer(distance * 1000) %>%
      st_transform(crs = 4326)
  })
  
  # Read in western_line parish list
  western_line_parishes <- read_excel('data/parishes/western_line_parishes.xlsx')
  
  # Create column called ref_code, by concatenating the "SE/" and variable parish_code, if there are 9 digits, add a 0 in front
  western_line_parishes$ref_code <- paste0('SE/', sprintf("%09d", western_line_parishes$parish_code))
  
  # Select only parish_name, parish_name_lower and ref_code
  western_line_parishes <- western_line_parishes %>% select(parish_name, parish_name_lower, ref_code)
  # Merge western_line_parishes with parishes
  parishes <- left_join(parishes, western_line_parishes %>% select(ref_code), by = "ref_code", suffix = c("", "_western_line"))
  parishes$treatment <- ifelse(!is.na(parishes$ref_code_western_line), 'western_line', 'control')
  
  # Create an object called western_line_parishes_geometry, which is a GeoDataFrame of the geometry of western_line parishes
  western_line_parishes_geometry <- left_join(western_line_parishes, parishes %>% select(ref_code, geometry), by = "ref_code")
  # Turn this into a GeoDataFrame
  western_line_parishes_geometry <- st_as_sf(western_line_parishes_geometry, crs = 4326, coords = c("longitude", "latitude"))
  
  # Calculate total power in each parish
  total_power <- stations_with_parish %>% group_by(ref_code) %>% summarise(amount_final = sum(amount_final, na.rm = TRUE))
  
  # Add these data to "parishes" and "control_parishes"
  parishes <- left_join(parishes, total_power, by = "ref_code")
  parishes <- left_join(parishes, num_power_stations, by = "ref_code")
  
  # In parishes, replace NaN with 0 in the outcome variables
  parishes$amount_final[is.na(parishes$amount_final)] <- 0
  parishes$num_power_stations[is.na(parishes$num_power_stations)] <- 0
  parishes$area_sqkm[is.na(parishes$area_sqkm)] <- 0
  
  control_parishes <- list()
  
  for (i in seq_along(distances)) {
    control_parishes[[distances[i]]] <- st_join(parishes, buffer_gdfs[[i]], join = st_intersects)
  }
  
  # In control parishes, generate treatment column = 0
  for (i in seq_along(distances)) {
    control_parishes[[distances[i]]]$treatment <- 0
  }
  
  # Find all parishes that touch a western_line parish
  if (exclude_touching) {
    touching_parishes <- st_intersection(parishes, western_line_parishes_geometry)
    # Drop ref_code_2 and rename ref_code_1 to ref_code
    touching_parishes <- touching_parishes %>% select(-ref_code_2) %>% rename(ref_code = ref_code_1)
    
    # Exclude these parishes from control groups
    for (i in seq_along(distances)) {
      control_parishes[[distances[i]]] <- control_parishes[[distances[i]]] %>% filter(!ref_code %in% touching_parishes$ref_code)
    }
  }
  
  # Create a DataFrame to store results
  results_df <- data.frame(Distance = numeric(), Source_Type = character(), Outcome_Var = character(), Exclude_Touching = logical(), Include_area = logical(), Include_population = logical(), Coefficient = numeric(), CI_lower = numeric(), CI_upper = numeric(), P_value = numeric(), F_statistic = numeric(), n_observations = numeric(), R_squared = numeric(), stringsAsFactors = FALSE)
  
  parishes_with_treatment <- parishes %>% filter(treatment == 'western_line')
  # Code treatment as 1
  parishes_with_treatment$treatment <- 1
  
  return(results_df)
}


```


## Census data from previous censuses for balance tests

First we read in the census data from [IPUMS](https://international.ipums.org/) the extract specified here:

#### Census Samples

| Sample       | Density | Note |
|--------------|---------|------|
| Sweden 1880  | 100.0%  |      |
| Sweden 1890  | 100.0%  |      |
| Sweden 1900  | 100.0%  |      |
| Sweden 1910  | 100.0%  |      |

#### Variables

| Type | Variable              | Label                                |
|------|------------------------|--------------------------------------|
| H    | COUNTRY               | Country                              |
| H    | YEAR                  | Year                                 |
| H    | SAMPLE                | IPUMS sample identifier              |
| H    | SERIAL                | Household serial number              |
| H    | HHWT                  | Household weight                     |
| H    | GEO1_SE1880_1910      | Sweden, County 1880 - 1910           |
| H    | PARSE                 | Sweden, Parish                       |
| P    | PERNUM                | Person number                        |
| P    | PERWT                 | Person weight                        |
| P    | AGE                   | Age                                  |
| P    | SEX                   | Sex                                  |
| P    | BIRTHYR               | Year of birth                        |
| P    | BPLSE1                | County of birth, Sweden              |
| P    | BPLSE2                | Parish of birth, Sweden              |
| P    | OCCISCO               | Occupation, ISCO general             |
| P    | OCCHISCO              | Occupation, HISCO classification     |
| P    | SE1880A_OCCSTRNG      | Occupation as transcribed            |
| P    | SE1890A_OCCSTRNG      | Occupation as transcribed            |
| P    | SE1900A_OCCSTRNG      | Occupation as transcribed            |
| P    | SE1910A_OCCSTRNG      | Occupation as transcribed            |

**Data Format**: `.dat (fixed-width text)`

**Structure**: Rectangular (person)

**Estimated Size**: 7147.2 MB


```{r}
#| eval: false

# Example data
df <- read_ipums_micro("path/ipums.xml")

df_clean <- df %>%
    janitor::clean_names()

df_short <- df_clean %>%
    # select the relevant columns
    select(year, contains("occ"), parse, bplse1, bplse2, sex, birthyr) %>%
    # remove labels to make files smaller
    zap_labels() %>%
    mutate(occstring = case_when(
        year == 1880 ~ se1880a_occstrng,
        year == 1890 ~ se1890a_occstrng,
        year == 1900 ~ se1900a_occstrng,
        year == 1910 ~ se1910a_occstrng,
        TRUE ~ "error"
    )) %>%
    select(-contains("se1")) %>%
    # tidy up occupational strings
    mutate(
        occstring = str_to_lower(occstring),
        occstring = str_squish(occstring)
    ) 

df_short

```

```{r}
#| eval: false
hisco_1930 <- read_dta("hisco_1930.dta")
```

## Information on parishes in the 1930 Swedish census

### Parish list

The digitization of the 1930 Swedish census was not complete when we got access to it from the National Archives of Sweden. As a result, we include a dataset `parishes_1930.dta` that contains the unique parish names and reference codes from the 1930 census that we use in our analysis. The dataset `parishes_1930.dta` contains the following variables:

- `parish_name`: The name of the parish.
- `parish_code`: The reference code of the parish.

An excerpt of the dataset is shown below:

```{r}
parishes_1930 <- read_dta("parishes_1930.dta")

parishes_1930 %>% head()
```


### Western Line Parishes

We created a binary variable for if a parish lay on the `western_line` between two large hydropower plants in Sweden. The variable `western_line` is a binary variable, taking the value 1 if the parish is on the western line and 0 otherwise. The variable `western_line` is constructed based on the electricity lines that were built between the hydropower plants in the early 20th century, and hand coded based on the geographical location of the parishes which intersected the main electricity line. The dataset we show below contains the names and reference codes of the parishes that are on the Western Line for which there is at least one individual born in that parish in the 1930 census.

The list of Western Line parishes is located at `western_line_parishes_1930.dta`, in the codebooks folder. The dataset `western_line_parishes_1930.dta` contains the following variables:

- `parish_code`: The reference code of the parish.
- `parish_name`: The name of the parish.

An excerpt of the dataset is shown below:

```{r}
df <- read_dta("hisco_1930.dta")

df %>% head()
```

To code this up in Stata, you can join the dataset with the main dataset using the `parish_code` variable:

```{stata}
#| eval: false

* Load the western line parishes dataset
use "western_line_parishes_1930.dta", clear

* Generate a variable to indicate that the parish is on the western line
gen western_line = 1

* Merge with the main dataset based on parish_code
merge 1:m parish_code using "swedish_census_1930.dta"

* Verify the merge
list parish_code _merge

* Drop the _merge variable
drop _merge

```


