---
title: "Replication for figures in the Paper Power and Progress"
format: html
---

### Setup

To replicate this analysis, you will need to install the following R packages:

- `tidyverse` for data manipulation and visualization.
- `showtext` to enable custom fonts in plots.
- `ggtext` for enhanced text rendering in ggplot2.
- `sf` for handling and visualizing spatial data.
- `here` to simplify file paths.
- `gghighlight` to highlight data in plots.
- `magick` for combining and manipulating images.
- `glue` for string interpolation in plot labels.

You can install these packages with the following command:

```{r}
#| eval: false
install.packages(c("tidyverse", "showtext", "ggtext", "sf", "here", "gghighlight", "magick", "glue"))
```

To reproduce the figures in the same style, ensure that Google fonts are accessible, as this script uses `IBM Plex Mono` and `Roboto`. You may need an active internet connection to download these fonts during the initial run.

```{r}
# Load packages
library(tidyverse)
library(showtext)
library(ggtext)
library(sf)
library(here)
library(magick)
library(glue)

# Set params
theme_set(theme_light())
font_add_google("IBM Plex Mono", "ibm")
font_add_google("Roboto", "roboto")

showtext_opts(dpi = 300)
showtext_auto(enable = TRUE)

# Set the working directory
# here::set_here("/Users/jonathanjayes/Dropbox/Electrification/replication")
# Check the working directory
# here::here()
```

### Figure 1

```{r}
# load the data
df <- read_rds(here::here("data", "figure-1.rds"))

tibble(
    year = c(1870, 1880, 1900, 1920, 1940, 1950),
    `Bottom (P0-P50)` = c(11, 17, 19, 16, 20, 21),
    `Middle (P50-P90)` = c(32, 32, 33, 31, 44, 47),
    `Top (P90-P100)` = c(57, 51, 48, 53, 36, 32)
) %>%
    pivot_longer(-year) %>%
    mutate(name = fct_relevel(name, "Bottom (P0-P50)", "Middle (P50-P90)", "Top (P90-P100)")) %>%
    mutate(name = fct_rev(name)) %>%
    ggplot() +
    geom_area(aes(year, value, fill = name), alpha = .6) +
    geom_line(aes(y = value, x = year, colour = "Rural Electrification Rate"), size = 1.1, data = df) +
    geom_vline(xintercept = 1921, linetype = "dashed", color = "#A6761D", size = 1.1) +
    geom_vline(xintercept = 1930, linetype = "solid", color = "#666666", size = 1.1) +
    annotate("text", x = 1921, y = 10, label = "Western Line Completed", hjust = -0.05, colour = "#A6761D") +
    annotate("text", x = 1930, y = 5, label = "Outcome Measured", hjust = -0.05, colour = "#666666") +
    scale_fill_brewer(palette = "Dark2") +
    scale_colour_manual(values = "midnightblue") +
    scale_y_continuous(labels = scales::percent_format(scale = 1), limits = c(0, NA)) +
    labs(
        x = NULL,
        y = "Share of income by group &\nShare of electrified rural households",
        title = NULL,
        fill = "Group in income distribution",
        colour = NULL
    ) +
    theme(
        panel.grid.minor.y = element_blank(),
        panel.grid.minor.x = element_blank(),
        plot.title = element_markdown(size = 32, family = "roboto"),
        plot.title.position = "plot",
        legend.position = "bottom",
        text = element_text(family = "ibm", size = 12),
        legend.box = "vertical"
    ) +
    guides(
        fill = guide_legend(nrow = 3, byrow = TRUE, title.position = "top", order = 2),
        colour = guide_legend(order = 1)
    )

ggsave(filename = here::here("output", "Figure-1.png"), device = "png", width = 290, height = 231, units = "mm", dpi = 300)
```

### Figure 2

```{r}
# Load data for Figure 2 (map of Western Line)
figure_2_data <- read_rds(here::here("data", "figure-2.rds"))

# Plot for the left panel of paper's Figure 2
plot1 <- ggplot() +
    geom_sf(data = figure_2_data$counties, colour = "grey70", fill = "#edf8fb") +
    geom_point(aes(geometry = geometry, colour = "Hydropower plants"),
        stat = "sf_coordinates",
        size = 10, shape = 18, alpha = .9,
        data = figure_2_data$hydro
    ) +
    geom_sf(data = figure_2_data$parishes_subset_treated_1930, colour = "grey70", aes(fill = "Treated parishes")) +
    geom_sf(data = figure_2_data$parishes_subset_control_1930, colour = "grey70", aes(fill = "Control parishes")) +
    geom_sf_text(aes(label = name, geometry=geometry), data = figure_2_data$hydro, stat = "sf_coordinates", size = 8, nudge_y = .35) +
    geom_sf(aes(colour = "Electricity lines\n1926"),
        data = figure_2_data$map_grid %>% filter(year == 1926), 
        linewidth = 1.5, alpha = .3, show.legend = "line"
    ) +
    scale_fill_manual(
        values = c("Treated parishes" = "#fc8d59", "Control parishes" = "#fef0d9"),
        guide = guide_legend(override.aes = list(linetype = "blank", shape = NA))
    ) +
    scale_color_manual(
        values = c("Electricity lines\n1926" = "#000080", "Hydropower plants" = "#d7301f"),
        guide = guide_legend(override.aes = list(linetype = c("solid", "blank"), shape = c(NA, 18)))
    ) +
    coord_sf(xlim = c(NA, 23), ylim = c(55.6, 65)) +
    theme_void() +
    labs(fill = NULL, colour = NULL) +
    theme(
        panel.background = element_rect(fill = "#DCE9ED"),
        legend.position = c(0.77, 0.11),
        text = element_text(family = "ibm", size = 14)
    )

ggsave(plot1, filename = here::here("output", "Figure-2_1.png"), device = "png", width = 290, height = 231, units = "mm", dpi = 300)

# Plot for the right panel of paper's Figure 2
plot2 <- ggplot() +
    geom_sf(data = figure_2_data$counties, colour = "grey70", fill = "#edf8fb") +
    geom_point(aes(geometry = geometry, colour = "Hydropower plants"),
        stat = "sf_coordinates",
        size = 10, shape = 18, alpha = .9,
        data = figure_2_data$hydro
    ) +
    geom_sf(data = figure_2_data$parishes_subset_treated, colour = "grey70", aes(fill = "Treated parishes")) +
    geom_sf(data = figure_2_data$parishes_subset_control, colour = "grey70", aes(fill = "Control parishes")) +
    geom_sf_text(aes(label = name, geometry = geometry), data = figure_2_data$hydro, stat = "sf_coordinates", size = 8, nudge_y = .35) +
    geom_sf(aes(colour = "Electricity lines\n1926"),
        data = figure_2_data$map_grid %>% filter(year == 1926), 
        linewidth = 1.5, alpha = .3, show.legend = "line"
    ) +
    scale_fill_manual(
        values = c("Treated parishes" = "#fc8d59", "Control parishes" = "#fef0d9"),
        guide = guide_legend(override.aes = list(linetype = "blank", shape = NA))
    ) +
    scale_color_manual(
        values = c("Electricity lines\n1926" = "#000080", "Hydropower plants" = "#d7301f"),
        guide = guide_legend(override.aes = list(linetype = c("solid", "blank"), shape = c(NA, 18)))
    ) +
    coord_sf(xlim = c(NA, 23), ylim = c(55.6, 65)) +
    theme_void() +
    labs(fill = NULL, colour = NULL) +
    theme(
        panel.background = element_rect(fill = "#DCE9ED"),
        legend.position = c(0.77, 0.11),
        text = element_text(family = "ibm", size = 14)
    )

ggsave(plot2, filename = here::here("output", "Figure-2_2.png"), device = "png", width = 290, height = 231, units = "mm", dpi = 300)

# Combine images
image1_mgk <- image_read(here::here("output", "Figure-2_1.png"))
image2_mgk <- image_read(here::here("output", "Figure-2_2.png"))
combined_image <- image_append(c(image2_mgk, image1_mgk), stack = FALSE)
image_write(combined_image, here::here("output", "Figure-2.png"))
```

### Figure 3

```{r}
df <- read_rds(here::here("data", "figure-3.rds"))

df %>%
    ggplot(aes(decile, coefficient)) +
    geom_hline(yintercept = 0, lty = 2) +
    geom_errorbar(aes(ymin = coefficient - 1.96 * standard_error, ymax = coefficient + 1.96 * standard_error), width = 2, colour = "#d95f02") +
    geom_line(colour = "#d95f02") +
    geom_point(cex = 3, colour = "#d95f02") +
    scale_x_continuous(labels = scales::percent_format(scale = 1), breaks = seq(10, 90, 10)) +
    labs(
        x = "Percentile of the income distribution",
        y = "Magnitude of regression coefficient"
    ) +
    theme(
        plot.title = element_markdown(size = 32, family = "roboto"),
        plot.title.position = "plot",
        legend.position = "none",
        text = element_text(family = "ibm", size = 18)
    )

ggsave(filename = here::here("output", "Figure-3.png"), device = "png", width = 290, height = 231, units = "mm", dpi = 300)
```

### Figure 4

```{r}
df <- read_rds(here::here("data", "figure-4.rds")) 

ggplot(df, aes(x = population, y = income_cumulative, colour = factor(current_parish_treated))) +
    geom_line() +
    geom_abline(slope = 1, color = "gray30", linetype = "dashed") +
    labs(x = "Cumulative Share of Population", y = "Cumulative Share of Income") +
    scale_colour_brewer(palette = "Dark2") +
    coord_cartesian() +
    guides(colour = guide_legend(reverse = TRUE)) +
    labs(colour = "Parish of residence 1930") +
    theme(legend.position = "bottom") +
    theme(
        panel.grid.minor.y = element_blank(),
        panel.grid.minor.x = element_blank(),
        plot.title = element_markdown(size = 32, family = "roboto"),
        plot.title.position = "plot",
        text = element_text(family = "ibm", size = 18)
    )

ggsave(filename = here::here("output", "Figure-4.png"), device = "png", width = 290, height = 290, units = "mm", dpi = 300)
```

### Figure 5

```{r}
df <- read_rds(here::here("data", "figure-5.rds"))

df %>%
    mutate(outcome_var = case_when(
        outcome_var == "Direct" ~ "Direct Electricity Jobs",
        outcome_var == "Indirect" ~ "Indirect Electricity Jobs"
    )) %>%
    filter(
        specification != "Birth Parish Treated + individual controls",
        year != 1910
    ) %>%
    ggplot(aes(x = year, y = b_birth_parish_treated, colour = outcome_var)) +
    geom_hline(yintercept = 0, linetype = 1, color = "black") +
    geom_rect(aes(xmin = 1909, xmax = 1922, ymin = -Inf, ymax = Inf), fill = "grey", alpha = 0.2, colour = NA) +
    geom_point() +
    geom_errorbar(aes(ymin = ymin, ymax = ymax), width = 0.2) +
    facet_wrap(~outcome_var, nrow = 2, scales = "free_y") +
    scale_colour_brewer(palette = "Dark2") +
    theme_minimal() +
    scale_y_continuous(labels = scales::number_format(scale = 100)) +
    theme(
        legend.position = "bottom",
        text = element_text(family = "ibm", size = 18)
    ) +
    labs(
        x = "Year",
        y = "LPM Estimate for Holding an Electricity Related Job\n(y-axis scaled by 100)",
        colour = "Job Type"
    )

ggsave(filename = here::here("output", "Figure-5.png"), device = "png", width = 290, height = 231, units = "mm", dpi = 300)
```

### Figure 6

```{r}
df <- read_rds(here::here("data", "figure-6.rds"))
    
df %>% 
    ggplot(aes(pct_share, napp_title, fill = Electricity_related)) +
    geom_col(position = "dodge") +
    geom_text(aes(label = pct_share_lab), hjust = -0.1) +
    scale_x_continuous(labels = scales::percent_format(), limits = c(NA, .2), breaks = c(0, .05, .1, .15)) +
    scale_fill_brewer(palette = "Dark2") +
    facet_grid(Electricity_related ~ treated, scales = "free_y") +
    labs(
        fill = "Electricity related job",
        y = NULL,
        x = "Share of total jobs"
    ) +
    theme(
        panel.grid.minor.x = element_blank(),
        panel.grid.minor.y = element_blank(),
        plot.title = element_markdown(size = 32, family = "roboto"),
        plot.subtitle = element_markdown(size = 18, family = "roboto"),
        plot.title.position = "plot",
        legend.position = "none",
        text = element_text(family = "ibm", size = 18)
    )

ggsave(filename = here::here("output", "Figure-6.png"), device = "png", width = 290, height = 231, units = "mm", dpi = 300)
```

### Figure 7

```{r}
df <- read_rds(here::here("data", "figure-7.rds"))

# Creating the coefficient plot for education interaction
# NOTE: The original script used 'new_data', which was not defined. Corrected to use 'df'.
ggplot(df, aes(y = variable, x = coefficient)) +
  geom_vline(xintercept = 0, color = "gray20", linetype = "dashed") +
  geom_point() +
  geom_errorbarh(aes(xmin = lower_ci, xmax = upper_ci), height = 0.2) +
  theme_minimal() +
  labs(x = "Magnitude of Regression Coefficient for\nVariable Interacted with Birth Parish (Treated)", y = "Variable Interacted with Birth Parish (Treated)") +
  theme(
    panel.grid.minor.y = element_blank(),
    panel.grid.minor.x = element_blank(),
    plot.title = element_markdown(size = 32, family = 'roboto'),
    plot.title.position = "plot",
    text = element_text(family = 'ibm', size = 18)
  )

ggsave(filename = here::here("output", "Figure-7.png"), device = "png", width = 290, height = 231, units = "mm", dpi = 300)
```

### Figure 8

```{r}
df <- read_rds(here::here("data", "figure-8.rds"))

ggplot(df, aes(x = decile, y = coefficient, color = Variable, lty = Variable, shape = Variable)) +
  geom_line() +
  geom_point(cex = 2, alpha = .8) +
  geom_errorbar(aes(ymin = coefficient - se, ymax = coefficient + se), width = 0.2) +
  theme_light() +
  scale_colour_brewer(palette = "Dark2") +
  geom_hline(yintercept = 0, lty = 2) +
  scale_x_continuous(labels = scales::percent_format(scale = 1), breaks = seq(10, 90, 10)) +
  theme(legend.position = "bottom") +
  labs(
    x = "Percentile of the income distribution",
    y = "Magnitude of regression coefficient (standardized)",
    color = "Variable",
    lty = "Variable",
    shape = "Variable"
  ) +
  theme(
    plot.title = element_markdown(size = 32, family = "roboto"),
    plot.title.position = "plot",
    legend.position = "bottom",
    text = element_text(family = "ibm", size = 18)
  )

ggsave(filename = here::here("output", "Figure-8.png"), device = "png", width = 290, height = 231, units = "mm", dpi = 300)
```

### Figure 9

```{r}
# This code replicates Figure 9 from the paper
# Data is created manually
counterfactual_data <- tibble::tribble(
  ~Decile, ~Period, ~Value,
  2, "1900-1910", 10,
  2, "1910-1920", 22,
  2, "1920-1930", 2,
  2, "1930-1940", 12,
  3, "1900-1910", 5,
  3, "1910-1920", 10,
  3, "1920-1930", 1,
  3, "1930-1940", 6,
  4, "1900-1910", 4,
  4, "1910-1920", 8,
  4, "1920-1930", 1,
  4, "1930-1940", 4,
  5, "1900-1910", 3,
  5, "1910-1920", 6,
  5, "1920-1930", 1,
  5, "1930-1940", 3,
  6, "1900-1910", 2.5,
  6, "1910-1920", 5.5,
  6, "1920-1930", 0.5,
  6, "1930-1940", 2.5,
  7, "1900-1910", 2,
  7, "1910-1920", 4,
  7, "1920-1930", 0.5,
  7, "1930-1940", 2,
  8, "1900-1910", 1.5,
  8, "1910-1920", 3.5,
  8, "1920-1930", 0.5,
  8, "1930-1940", 1.5,
  9, "1900-1910", 1,
  9, "1910-1920", 3,
  9, "1920-1930", 0.5,
  9, "1930-1940", 1.5,
  10, "1900-1910", 1,
  10, "1910-1920", 2.5,
  10, "1920-1930", 0.5,
  10, "1930-1940", 1
)

counterfactual_data %>%
  mutate(Period = factor(Period, levels = c("1930-1940", "1920-1930", "1910-1920", "1900-1910"))) %>%
  ggplot(aes(x = factor(Decile), y = Value, fill = Period)) +
  geom_col() +
  scale_fill_manual(
    values = c("1900-1910" = "#0072B2", "1910-1920" = "#D55E00", "1920-1930" = "#009E73", "1930-1940" = "#F0E442"),
    name = NULL
  ) +
  labs(
    x = "Decile",
    y = "Counterfactual percentage change in income (%)"
  ) +
  theme_light() +
  theme(
    legend.position = "top",
    text = element_text(family = "ibm", size = 18)
  )

ggsave(filename = here::here("output", "Figure-9.png"), device = "png", width = 290, height = 231, units = "mm", dpi = 300)

```

---
### Appendix Figures
---

### Appendix Figure A.10: Grid Roll-out

```{r}
df <- read_rds(here::here("data", "figure-10.rds")) # Formerly figure-11.rds

plot_figure_a10 <- function(year_to_plot) {
    
    ggplot() +
        geom_sf(data = df$counties, colour = "grey70", fill = "#edf8fb") +
        geom_point(aes(geometry = geometry, colour = "Hydropower plants"),
                             stat = "sf_coordinates",
                             size = 10, shape = 18, alpha = .9,
                             data = df$hydro
        ) +
        geom_sf_text(aes(label = name, geometry = geometry), stat = "sf_coordinates", data = df$hydro, size = 8, nudge_y = .35) +
        geom_sf(aes(colour = glue("Electricity lines\n{year_to_plot}"), geometry = geometry),
                        data = df$map_grid %>% filter(year == year_to_plot), 
                        linewidth = 1.5, alpha = .5, show.legend = "line"
        ) +
        scale_color_manual(
            values = c(glue("Electricity lines\n{year_to_plot}") = "#000080", "Hydropower plants" = "#d7301f"),
            guide = guide_legend(override.aes = list(linetype = c("solid", "blank"), shape = c(NA, 18)))
        ) +
        coord_sf(xlim = c(NA, 23), ylim = c(55.6, 65)) +
        theme_void() +
        labs(colour = NULL) +
        theme(
            panel.background = element_rect(fill = "#DCE9ED"),
            legend.position = c(0.77, 0.11),
            text = element_text(family = "ibm", size = 15)
        )
}

# Example usage for 1926 from paper's Figure 10
plot_figure_a10(1926)

ggsave(filename = here::here("output", "Appendix-Figure-A10.png"), device = "png", width = 290, height = 231, units = "mm", dpi = 300)
```

### Appendix Figure A.11: Propensity Score Comparison

```{r}
df <- read_rds(here::here("data", "figure-11.rds")) # Formerly figure-12.rds

df %>%
    ggplot(aes(x = value, y = treatment, colour = treatment)) +
    geom_point(alpha = .7) +
    facet_wrap(~period) +
    scale_colour_brewer(palette = "Dark2") +
    theme(legend.position = "none") +
    labs(x = "Propensity Score", y = NULL)

ggsave(filename = here::here("output", "Appendix-Figure-A11.png"), device = "png", width = 290, height = 231, units = "mm", dpi = 300)
```

### Appendix Figure A.12: Marginal effect of age

```{r}
df <- read_rds(here::here("data", "figure-12.rds")) # Formerly figure-13.rds

ggplot(df, aes(x = age, y = estimate)) +
    geom_errorbar(aes(ymin = conf.low, ymax = conf.high)) +
    geom_point(colour = "#1A9973", cex = 2) +
    geom_hline(yintercept = 0, lty = 2) +
    theme_minimal() +
    scale_x_continuous(breaks = c(15, 30, 45, 60)) +
    labs(x = "Age", y = "Marginal effect of age on Western Line parish coefficient") +
    theme(
        panel.grid.minor.y = element_blank(),
        panel.grid.minor.x = element_blank(),
        plot.title = element_markdown(size = 32, family = "roboto"),
        plot.title.position = "plot",
        legend.position = "none",
        text = element_text(family = "ibm", size = 18)
    )

ggsave(filename = here::here("output", "Appendix-Figure-A12.png"), device = "png", width = 290, height = 231, units = "mm", dpi = 300)
```

### Appendix Figure A.13: Conditional vs. Residualized Quantile Regression

```{r}
df <- read_rds(here::here("data", "figure-13.rds")) # Formerly figure-15.rds

df %>%
    mutate(estimator = fct_relevel(estimator, "Residualized QR", "Conditional QR")) %>%
    ggplot(aes(decile, coefficient, colour = estimator, lty = estimator, shape = estimator)) +
    geom_hline(yintercept = 0, lty = 2) +
    geom_errorbar(aes(ymin = lower, ymax = upper), width = 2) +
    geom_line() +
    geom_point(cex = 3) +
    scale_colour_brewer(palette = "Dark2") +
    scale_x_continuous(labels = scales::percent_format(scale = 1), breaks = seq(10, 90, 10)) +
    scale_linetype_manual(values = c("Conditional QR" = "solid", "Residualized QR" = "dashed")) +
    theme(legend.position = "bottom") +
    labs(
        x = "Percentile of the income distribution",
        y = "Magnitude of regression coefficient",
        colour = "Estimator",
        shape = "Estimator",
        lty = "Estimator"
    ) +
    theme(
        plot.title = element_markdown(size = 32, family = "roboto"),
        plot.title.position = "plot",
        legend.position = "bottom",
        text = element_text(family = "ibm", size = 18)
    )

ggsave(filename = here::here("output", "Appendix-Figure-A13.png"), device = "png", width = 290, height = 231, units = "mm", dpi = 300)
```

### Appendix Figure A.14: Union density interaction (1910 & 1930)

```{r}
df <- read_rds(here::here("data", "figure-14.rds")) # Formerly figure-14.rds

ggplot(df, aes(x = decile, y = coefficient, color = Variable, lty = Variable, shape = Variable)) +
    geom_line() +
    geom_point(cex = 2, alpha = .8) +
    geom_errorbar(aes(ymin = coefficient - se, ymax = coefficient + se), width = 0.2) +
    facet_grid(year ~ .) +
    theme_light() +
    scale_colour_brewer(palette = "Dark2") +
    geom_hline(yintercept = 0, lty = 2) +
    scale_x_continuous(labels = scales::percent_format(scale = 1), breaks = seq(10, 90, 10)) +
    theme(legend.position = "bottom") +
    labs(
        x = "Percentile of the income distribution",
        y = "Magnitude of regression coefficient"
    ) +
    theme(
        plot.title = element_markdown(size = 32, family = "roboto"),
        plot.title.position = "plot",
        legend.position = "bottom",
        text = element_text(family = "ibm", size = 18)
    )

ggsave(filename = here::here("output", "Appendix-Figure-A14.png"), device = "png", width = 290, height = 231, units = "mm", dpi = 300)
